Relationships in the extremes and their influence on competition and coexistence

### 2021-09-30 paper/Paper.Rnw
fixed mu's in fig 2 caption and spelling and added Sexpr{rho} to fig 1 caption

\begin{figure}
\includegraphics[width=.25\textwidth]{../results_figs/fig1_vertical.pdf}
\caption{Simple example of bivariate random variables with standard normal marginals showing
alternative patterns of association in the tails. A) is ``left-tail 
associated,'' B) is symmetric, and C) is ``right-tail associated'' noise. All three cases have 
the same Pearson correlation, $\rho$, up to sampling variation, yet show very different 
patterns of association. We used asymmetrically tail associated cases which are perfectly 
correlated in the lower (A)  or upper (C) halves of the distributions, though it is also 
possible to generate asymmetrically tail associated noise with imperfect correlations in 
both halves of the distributions \citep{Ghosh_2020_AER,Nelsen2006}. The variables $(b_1,b_2)$ described in
the Introduction and Methods were obtained by generating 
noise as pictured and then multipling by $\sigma$ and adding $(\mu_1,\mu_2)$.\label{fig:noise}}
\end{figure}
%***DAN: Dan has edited this caption at least once.

\begin{figure}
\includegraphics[width=.75\textwidth]{../results_figs/fig2.pdf}
\caption{Our simulated lottery-model populations influenced by asymmetric tail associations (ATAs) recovered from low densities less often and 
spent more time in a state of one-species dominance, demonstrating that ATAs can influence aspects of 
species coexistence. Panel A shows species 1, from the lottery 
model subject to environmental noise with symmetric tail associations (Fig. \ref{fig:noise}B). Panel B shows species 1 from the 
same model with the same parameters, but subject to noise with left-tail association (Fig. \ref{fig:noise}A).
Both simulations were run for \Sexpr{M} time steps, though only the first $1000$ are shown. Periods of 
``noticeable coexistence'' of the species, defined to be periods when both species each comprise at least $5\%$ of 
total abundance, are marked in grey. The average duration of periods of noticiable coexistence 
was \Sexpr{fig2$coSym} time steps for panel A and \Sexpr{fig2$coATA} for panel B, over the full length of the simulations; 
whereas the average duration of periods of
one-species dominance was \Sexpr{fig2$domSym} time steps for panel A and \Sexpr{fig2$domATA} for B. Thus, while periods of 
noticeable coexistence were of similar duration for the two scenarios, periods of one-species dominance were much longer,
on average, for the ATA case. Parameters were $\mu_1-\mu_2 = \Sexpr{fig2$params[1]}$, 
%***DAN: Jasmin, at this point of the ms, the reader does not know that only the difference mu1-mu2 matters, so this
%will confuse them. Please replace the above specfication of mu1-mu2 by a specification of which values of both
%mu1 and mu2 were actually used for this simulation. 
$\delta = \Sexpr{fig2$params[2]}$, and $\sigma = \Sexpr{fig2$params[3]}$. See Introduction for model details. 
\label{fig:pedagog}}
\end{figure}
%***DAN: Dan has edited this caption

\begin{figure}
\includegraphics[width=\textwidth]{../results_figs/fig3.pdf}
\caption{The contribution of ATAs to species coexistence is often meaningfully large compared to 
other components of the MCT decomposition. All panels show the IGR and the decomposition of
the IGR provided by the extended MCT, equation (\ref{eq:full_decomp}). For each panel, 
$\delta$ and $\mu_1-\mu_2$ 
are fixed at values specified in the margins, and the IGR and components of the decomposition
are plotted against $\sigma$. The contribution of ATAs (red line) frequently 
has magnitude comparable to other components, indicating that ATAs can contribute meaningfully
to species coexistence or failure to coexist. In some instances, the IGR line (orange) falls 
betweed the ATA line (red)
and the $x$ axis, indicating that, for those parameter values, ATAs make the difference between species
coexistence and exclusion of the weaker competitor. Standard errors of plotted quantities 
were never greater than \Sexpr{round(se3,5)}, so are not displayed. This plot uses $q_{ij}=1$. An 
analogous plot (SI Fig. \ref{SI-fig:fig3_alt}) was generated 
using $q_{ij}=\exp(\mu_i)/[(1-\delta)\exp(\mu_j)+\delta\exp(\mu_i)]$. Though results differed
in some details, main conclusions were the same. \label{fig:fig3}}
\end{figure}
%***DAN: Dan has edited this caption

>>>>

\begin{figure}
\includegraphics[width=.25\textwidth]{../results_figs/fig1_vertical.pdf}
\caption{Simple example of bivariate random variables with standard normal marginals showing
alternative patterns of association in the tails. A) is ``left-tail 
associated,'' B) is symmetric, and C) is ``right-tail associated'' noise. All three cases have 
the same Pearson correlation, $\rho = \Sexpr{rho}$, up to sampling variation, yet show very different 
patterns of association. We used asymmetrically tail associated cases which are perfectly 
correlated in the lower (A)  or upper (C) halves of the distributions, though it is also 
possible to generate asymmetrically tail associated noise with imperfect correlations in 
both halves of the distributions \citep{Ghosh_2020_AER,Nelsen2006}. The variables $(b_1,b_2)$ described in
the Introduction and Methods were obtained by generating 
noise as pictured and then multipling by $\sigma$ and adding $(\mu_1,\mu_2)$.\label{fig:noise}}
\end{figure}
%***DAN: Dan has edited this caption at least once.

\begin{figure}
\includegraphics[width=.75\textwidth]{../results_figs/fig2.pdf}
\caption{Our simulated lottery-model populations influenced by asymmetric tail associations (ATAs) recovered from low densities less often and 
spent more time in a state of one-species dominance, demonstrating that ATAs can influence aspects of 
species coexistence. Panel A shows species 1, from the lottery 
model subject to environmental noise with symmetric tail associations (Fig. \ref{fig:noise}B). Panel B shows species 1 from the 
same model with the same parameters, but subject to noise with left-tail association (Fig. \ref{fig:noise}A).
Both simulations were run for \Sexpr{M} time steps, though only the first $1000$ are shown. Periods of 
``noticeable coexistence'' of the species, defined to be periods when both species each comprise at least $5\%$ of 
total abundance, are marked in grey. The average duration of periods of noticeable coexistence 
was \Sexpr{fig2$coSym} time steps for panel A and \Sexpr{fig2$coATA} for panel B, over the full length of the simulations; 
whereas the average duration of periods of
one-species dominance was \Sexpr{fig2$domSym} time steps for panel A and \Sexpr{fig2$domATA} for B. Thus, while periods of 
noticeable coexistence were of similar duration for the two scenarios, periods of one-species dominance were much longer,
on average, for the ATA case. Parameters were $\mu_1 = \Sexpr{fig2$params[1]} + \Sexpr{fig2$params[1]}$ and $\mu_2 = \Sexpr{-(fig2$params[1])}$, $\delta = \Sexpr{fig2$params[2]}$, and $\sigma = \Sexpr{fig2$params[3]}$. See Introduction for model details. 
\label{fig:pedagog}}
\end{figure}
%***DAN: Dan has edited this caption

\begin{figure}
\includegraphics[width=\textwidth]{../results_figs/fig3.pdf}
\caption{The contribution of ATAs to species coexistence is often meaningfully large compared to 
other components of the MCT decomposition. All panels show the IGR and the decomposition of
the IGR provided by the extended MCT, equation (\ref{eq:full_decomp}). For each panel, 
$\delta$ and $\mu_1-\mu_2$ 
are fixed at values specified in the margins, and the IGR and components of the decomposition
are plotted against $\sigma$. The contribution of ATAs (red line) frequently 
has magnitude comparable to other components, indicating that ATAs can contribute meaningfully
to species coexistence or failure to coexist. In some instances, the IGR line (orange) falls 
between the ATA line (red)
and the $x$ axis, indicating that, for those parameter values, ATAs make the difference between species
coexistence and exclusion of the weaker competitor. Standard errors of plotted quantities 
were never greater than \Sexpr{round(se3,5)}, so are not displayed. This plot uses $q_{ij}=1$. An 
analogous plot (SI Fig. \ref{SI-fig:fig3_alt}) was generated 
using $q_{ij}=\exp(\mu_i)/[(1-\delta)\exp(\mu_j)+\delta\exp(\mu_i)]$. Though results differed
in some details, main conclusions were the same. \label{fig:fig3}}
\end{figure}
%***DAN: Dan has edited this caption
